---
title: "Group17_code"
output: html_document
date: "2024-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment=NA,attr.source='.numberLines')
rm(list=ls())
library(readr)
library(RSQLite)
library(dplyr)
library(DBI)
```

```{r}
all_files <- list.files("data_upload/")
all_files
```
```{r block2,attr.source='.numberLines'}
prefix <- "hi_"
suffix <- "_dataset.csv"

all_files <- gsub("hi_","",all_files)
all_files <- gsub("_dataset.csv","",all_files)
all_files

```

# Looping through - for 
Check number of rows and columns

```{r loop,message=FALSE,warning=FALSE,attr.source='.numberLines'}
all_files <- list.files("data_upload/")

for (variable in all_files) {
  this_filepath <- paste0("data_upload/",variable)
  this_file_contents <- readr::read_csv(this_filepath)
  
  number_of_rows <- nrow(this_file_contents)
  number_of_columns <- ncol(this_file_contents)
  
  print(paste0("The file: ",variable,
              " has: ",
              format(number_of_rows,big.mark = ","),
              " rows and ",
              number_of_columns," columns"))
}
```

#Check the data structure
```{r loop,message=FALSE,warning=FALSE,attr.source='.numberLines'}
all_files <- list.files("data_upload/")

for (variable in all_files) {
  this_filepath <- paste0("data_upload/",variable)
  this_file_contents <- readr::read_csv(this_filepath)
  data_structure<-str(this_file_contents)
  
  print(paste0(data_structure,
               "The file: ",variable,
              " has above data structure"))
}
```
#Check for NULL values
```{r loop,message=FALSE,warning=FALSE,attr.source='.numberLines'}
all_files <- list.files("data_upload/")

for (variable in all_files) {
  this_filepath <- paste0("data_upload/",variable)
  this_file_contents <- readr::read_csv(this_filepath)
  null<-sum(is.na(this_file_contents))
  
  print(paste0("The file: ",variable,
              " has a total of ", null,
              " NULL values"))
}
```
#Check that each primary key is unique in each table
```{r loop,message=FALSE,warning=FALSE,attr.source='.numberLines'}
all_files <- list.files("data_upload/")

for (variable in all_files) {
  this_filepath <- paste0("data_upload/",variable)
  this_file_contents <- readr::read_csv(this_filepath)
  hi <- nrow(unique(this_file_contents[,1]))== nrow(this_file_contents)
  
  print(paste0("The file: ",variable,
              " has unique primary key ",
              hi," columns"))
}
```

# Load Files in an sqlite database 
```{r loadsqlite,warning=FALSE,error=FALSE,message=FALSE,attr.source='.numberLines'}
#setup the connection
connection <- RSQLite::dbConnect(RSQLite::SQLite(),"hi_import.db")
```



## Connect and Setting options in the SQLite terminal
* Let us connect to the ``hi_import.db`` database in the folder.
```{bash, eval=FALSE}
sqlite3 hi_import.db 
```

```{sql, eval=FALSE}
.headers on 
.mode column 
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_transactions_dataset.csv transactions
```
1. product_category
```{sql, eval=FALSE}
-- product_category
CREATE TABLE "product_category" ( 
  category_name VARCHAR(50) PRIMARY KEY,
  parent_category_id INT NULL
);
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_productcategory_dataset.csv "product_category"
```

```{sql connection=connection}
SELECT * FROM "product_category";
```

2. promotion
```{sql, eval=FALSE}
-- promotion
CREATE TABLE "promotion" ( 
  promo_code INT PRIMARY KEY, 
  percentage_discount NUMERIC NOT NULL
);
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_promotion_dataset.csv "promotion"
```

```{sql connection=connection}
SELECT * FROM "promotion";
```

3. supplier
```{sql, eval=FALSE}
-- supplier
CREATE TABLE supplier ( 
  supplier_id INT PRIMARY KEY, 
  supplier_name CHAR NOT NULL,
  supplier_phone INT NOT NULL,
  supplier_email VARCHAR(50) NOT NULL,
  supplier_building INT NOT NULL,
  supplier_street VARCHAR(50) NOT NULL,
  supplier_city VARCHAR(50) NOT NULL,
  supplier_region VARCHAR(50) NULL,
  supplier_postcode VARCHAR(50) NOT NULL
) ; 
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_supplier_dataset.csv "supplier"
```

```{sql connection=connection}
SELECT * FROM "supplier";
```

4. customer
```{sql, eval=FALSE}
-- customer
CREATE TABLE "customer" ( 
  customer_id INT PRIMARY KEY, 
  customer_firstname VARCHAR(50) NOT NULL,
  customer_lastname VARCHAR(50) NOT NULL,
  customer_title VARCHAR(25) NOT NULL, 
  customer_phone VARCHAR(50) NOT NULL,
  customer_email VARCHAR(50) NOT NULL,
  customer_membership TEXT NOT NULL, 
  delivery_fee NUMERIC NOT NULL,
  customer_building INT NOT NULL,
  customer_street VARCHAR(50) NOT NULL, 
  customer_city VARCHAR(50) NOT NULL, 
  customer_region VARCHAR(50) NULL, 
  customer_postcode VARCHAR(50) NOT NULL, 
  promo_code INT,
  FOREIGN KEY (promo_code) REFERENCES "promotion"(promo_code)
) ; 
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_customer_dataset.csv "customer"
```

```{sql connection=connection}
SELECT * FROM "customer";
```

5. delivery
```{sql, eval=FALSE}
-- delivery
CREATE TABLE "delivery" ( 
  tracking_number INT PRIMARY KEY, 
  shipment_method VARCHAR(50) NOT NULL,
  tracking_status VARCHAR(50) NOT NULL,
  estimated_delivery_date DATE NOT NULL,
  estimated_delivery_time TIME NOT NULL,
  actual_delivery_date DATE NULL, 
  actual_delivery_time TIME NULL,
  delivery_instructions VARCHAR(125) NOT NULL,
  trans_id INT,
  FOREIGN KEY (trans_id) REFERENCES "transaction"(trans_id)
); 
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_delivery_dataset.csv "delivery"
```

```{sql connection=connection}
SELECT * FROM "delivery";
```

6. product
```{sql, eval=FALSE}
-- product
CREATE TABLE "product" ( 
  product_id INT PRIMARY KEY, 
  product_name VARCHAR(25) NOT NULL,
  product_weight NUMERIC NOT NULL,
  product_length NUMERIC NOT NULL,
  product_height NUMERIC NOT NULL,
  product_width NUMERIC NOT NULL,
  product_price NUMERIC NOT NULL,
  supplier_id INT,
  category_name VARCHAR(50),
  FOREIGN KEY (supplier_id) REFERENCES "supplier"(supplier_id),
  FOREIGN KEY (category_name) REFERENCES "product_category"(category_name)
) ; 
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_product_dataset.csv "product"
```

```{sql connection=connection}
SELECT * FROM "product";
```

7. order
```{sql connection=connection}
-- order
CREATE TABLE "order" (
  customer_id INT,
  order_id INT,
  product_id INT,
  product_qty INT NOT NULL,
  order_date DATE NOT NULL, 
  order_time TIME NOT NULL,
  PRIMARY KEY (customer_id, order_id, product_id),
  FOREIGN KEY (customer_id) REFERENCES "customer"(customer_id),
  FOREIGN KEY (product_id) REFERENCES "product"(product_id)
) ; 
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_orders_dataset.csv "order"
```

```{sql connection=connection}
SELECT * FROM "order";
```

8. transaction
```{sql connection=connection}
-- transaction
CREATE TABLE "transaction" (
  trans_id INT PRIMARY KEY,
  order_id INT,
  trans_date DATE NOT NULL,
  trans_time TIME NOT NULL,
  FOREIGN KEY (order_id) REFERENCES "order"(order_id)
);
```

```{sql connection=connection}
.mode csv
.import --skip 1 data_upload/hi_transactions_dataset.csv "transaction"
```

```{sql connection=connection}
SELECT * FROM "transaction";
```

Store natively it in R
```{r}
customer <- dbReadTable(connection, "customer")
delivery <- dbReadTable(connection, "delivery")
order <- dbReadTable(connection, "order")
product <- dbReadTable(connection, "product")
product_category <- dbReadTable(connection, "product_category")
promotion <- dbReadTable(connection, "promotion")
supplier <- dbReadTable(connection, "supplier")
transaction <- dbReadTable(connection, "transaction")
```


## Now normalize to 3NF
For customer
1. customer_memebership
```{sql connection=connection}
CREATE TABLE customer_memebership AS
SELECT customer_id, customer_membership
FROM "customer";
```

```{sql connection=connection}
SELECT * FROM customer_memebership
```

2. delivery_fee_info
```{sql connection=connection}
CREATE TABLE delivery_fee_info AS
SELECT customer_membership, delivery_fee
FROM "customer";
```

```{sql connection=connection}
SELECT * FROM delivery_fee_info
```

3. customer_basic_info
```{sql connection=connection}
CREATE TABLE customer_basic_info AS
SELECT customer_id, customer_firstname, customer_lastname, customer_title, customer_phone, customer_email, customer_building, customer_street,  customer_city,  customer_region,  customer_postcode, promo_code
FROM "customer";
```

```{sql connection=connection}
SELECT * FROM customer_basic_info
```

For order
1. order_datetime
```{sql connection=connection}
CREATE TABLE order_datetime AS
SELECT DISTINCT customer_id, order_id, order_date, order_time
FROM "order";
```

```{sql connection=connection}
SELECT * FROM order_datetime
```

2. order_products_info
```{sql connection=connection}
CREATE TABLE order_products_info AS
SELECT customer_id, order_id, product_id, product_qty
FROM "order";
```

```{sql connection=connection}
SELECT * FROM order_products_info
```

For Delivery
1. tracking_number
```{sql connection=connection}
CREATE TABLE tracking_number AS
SELECT tracking_number, delivery_instructions, trans_id
FROM "delivery";
```

```{sql connection=connection}
SELECT * FROM tracking_number
```

2. estimated_delivery_date
```{sql connection=connection}
CREATE TABLE estimated_delivery_date AS
SELECT tracking_number, shipment_method, estimated_delivery_date, estimated_delivery_time
FROM "delivery";
```

```{sql connection=connection}
SELECT * FROM estimated_delivery_date
```

3. actual_delivery_date
```{sql connection=connection}
CREATE TABLE actual_delivery_date AS
SELECT tracking_number, tracking_status, actual_delivery_date, actual_delivery_time
FROM "delivery";
```

```{sql connection=connection}
SELECT * FROM actual_delivery_date
```


```{sql connection=connection}
PRAGMA table_info(actual_delivery_date);
```

Store natively it in R
```{r}
customer_memebership <- dbReadTable(connection, "customer_memebership")
delivery_fee_info <- dbReadTable(connection, "delivery_fee_info")
customer_basic_info <- dbReadTable(connection, "customer_basic_info")

order_datetime <- dbReadTable(connection, "order_datetime")
order_products_info <- dbReadTable(connection, "order_products_info")

tracking_number <- dbReadTable(connection, "tracking_number")
estimated_delivery_date <- dbReadTable(connection, "estimated_delivery_date")
actual_delivery_date <- dbReadTable(connection, "actual_delivery_date")
```

```{r listtables}
RSQLite::dbListTables(connection)
```

```{r disconnect}
RSQLite::dbDisconnect(connection)
```
